<!--- Begin styling code. --->
<style type="text/css">
/* Whole document: */
body{
  font-family: "Palatino Linotype", "Book Antiqua", Palatino, serif;
  font-size: 12pt;
}
h1.title {
  font-size: 38px;
  text-align: center;
}
h4.author {
  font-size: 18px;
  text-align: center;
}
h4.date {
  font-size: 18px;
  text-align: center;
}
</style>
<!--- End styling code. --->

---
title: "Short Lab 3"
author: "Sylvia Du"
date: "2/3/2020"
output: html_document
---

As usual, all code below should follow the style guidelines from the lecture slides.

## Part 1. Read in text data

For this short lab, we will be using Project Gutenberg’s The Complete Works of William Shakespeare. 
Use the command `read_lines()` to read the text available at
"https://www.gutenberg.org/files/100/100-0.txt".
Make sure to store the text as a variable.

```{r, eval = TRUE, echo = TRUE, message = FALSE}
# import libraries that are needed in this shortlab
library(tidyverse)
library(lubridate)

# store the data from the website into a variable
my_data <- read_lines("https://www.gutenberg.org/files/100/100-0.txt")
```


**1a.** Print the first 5 lines.

```{r, eval = TRUE, echo = TRUE, message = FALSE}
# to print out first 5 elements (lines) of data
my_data %>%
  head(5)
```

**1b.** Print the total number of lines.

```{r, eval = TRUE, echo = TRUE, message = FALSE}
# print the number of elements (lines) of the data
length(my_data)
```

**1c.** Remove all empty lines, then print the total number of lines.

(*Hint: to remove empty elements from a string vector x, you could use* `x <- x[x != ""]`)

```{r, eval = TRUE, echo = TRUE, message = FALSE}
# clear all empty elements from data and print the length of data
my_data <- my_data[my_data != ""]
length(my_data)
```


## Part 2. Regular expressions

**2a.** Use a regular expression with `str_count()` to count how much punctuation is in this text file, in total.

```{r, eval = TRUE, echo = TRUE, message = FALSE}
# print the sum of all punctuations in the data
sum(str_count(my_data, "[,.?';:!#*-`]"))
```

**2b.** Use a regular expression with `str_detect()` to count how many lines contain *either* the string "Romeo" or "Juliet".

```{r, eval = TRUE, echo = TRUE, message = FALSE}
# count the number of elements containing Romeo or Juliet
str_detect(my_data, "Romeo|Juliet") %>% 
  as.numeric %>%
  sum()
```


## Part 3. String Manipulation

**3a.** Use `str_c()` to collapse the Shakespeare string vector into one large string. (Don't try to print it!)

```{r, eval = TRUE, echo = TRUE, message = FALSE}
# collapse the data into a single string 
my_long <- str_c(my_data, collapse = " ")
```

**3b.** Use `str_split()` to separate your string into words.

(*Hint: you might get a list of length 1 that you have to convert to a vector. You could do this by using something like *`x <- unlist(x)` *or* `x <- x[[1]]`)

```{r, eval = TRUE, echo = TRUE, message = FALSE}
# split the string into words stored in a vector of strings
my_words <- unlist(str_split(my_long, " "))
```

**3c.** Use a combination of `table()` and `sort(..., decreasing = TRUE)` argument to get a count of the unique words in Shakespeare's complete works and print out the 10 most common words.

```{r, eval = TRUE, echo = TRUE, message = FALSE}
# count the number of each unique word and sort the by most used, print first 10
my_table <- sort(table(my_words), decreasing = TRUE)
my_table %>%
  head(10)
```


## Part 4. Factors

**4a.** Create an object that is a factor vector with 4 levels, where each of these levels is observed at least once.

```{r, eval = TRUE, echo = TRUE, message = FALSE}
# construct my factor with levels a, b, c, d and each observed once
my_factor <- factor(c("a", "b", "c", "d"), levels = c("a", "b", "c", "d"))
```

**4b.** Collapse two of your factor levels together into a new level "x".

```{r, eval = TRUE, echo = TRUE, message = FALSE}
# collapsee levels a and b into new level x
my_factor <- fct_collapse(my_factor, x = c("a", "b"))
```

**4c.** Add a new, empty level to your factor and print out the vector.

```{r, eval = TRUE, echo = TRUE, message = FALSE}
# add a new empty level e to my_factor and print out my_factor
my_factor <- fct_expand(my_factor, "e")
my_factor
```

**4d.** Remove this empty level from your factor and print out the vector.

```{r, eval = TRUE, echo = TRUE, message = FALSE}
# drop empty level e from my_factor and print my_factor
my_factor <- fct_drop(my_factor, only = "e")
my_factor
```


## Part 5. Dates

**5a.** Create a date-time object in R, with both a date and a time.

```{r, eval = TRUE, echo = TRUE, message = FALSE}
# creat a date-time object x with date of today and time of (approximately) now
x <- mdy_hms("Feb 1st 2020, 14:40:00")
```

**5b.** Extract the date from your object.

```{r, eval = TRUE, echo = TRUE, message = FALSE}
# create a new variable d equals to the date of x
d <- date(x)
```

**5c.** Extract the month from your object.

```{r, eval = TRUE, echo = TRUE, message = FALSE}
# create a new variable m equals to the month of x
m <- month(x)
```

**5d.** Change the hour of your object, then extract the hour from your object.

```{r, eval = TRUE, echo = TRUE, message = FALSE}
# change the hour of x into 20 and create a new variable h equals to the hour 
# of x
hour(x) <- 20
h <- hour(x)
```




