<!--- Please do not touch the code below! It is for styling. --->
<style type="text/css">
/* Whole document: */
body{
  font-family: "Palatino Linotype", "Book Antiqua", Palatino, serif;
  font-size: 12pt;
}
h1.title {
  font-size: 38px;
  text-align: center;
}
h4.author {
  font-size: 18px;
  text-align: center;
}
h4.date {
  font-size: 18px;
  text-align: center;
}
</style>
<!--- Begin here! Fill in text and R code for each of these questions. --->

---
title: "Short Lab 2"
author: "Sylvia Du"
date: "1/15/2020"
output: html_document
---

As usual, all code below should follow the style guidelines from the lecture slides.

### Part A. Control Flow

1. 
    a. Write a function that returns a Boolean indicating whether the input is numeric or not.
    b. Evaluate your function with input that returns `TRUE`
    c. Evaluate your function with input that returns `FALSE`

#a
```{r eval = TRUE, echo = TRUE}
# function: check_numeric, checks if the input is numeric and returns a Boolean
#           indicating the result.
# input: x
# output: Boolean
check_numeric <- function(x) {
    return(is.numeric(x))
}
```

#b
```{r eval = TRUE, echo = TRUE}
# let input be a number and check if it is numeric. if returns TRUE, the 
# function works.
a <- 3
check_numeric(a)
```

#c
```{r eval = TRUE, echo = TRUE}
# let input be a string and check if it is numeric. if returns FALSE, the 
# function works.
b <- "cat"
check_numeric(b)
```

2. 
    a. Write a function returns the absolute value of the input. This function should check to make sure the input is numeric. 
If the input is not numeric, it should `stop` with an informative error message.
    b. Evaluate your function with positive numeric input.
    c. Evaluate your function with negative numeric input.
    d. Trigger your custom error message. Note: You will need to add the parameter 
`error = TRUE` to your R Markdown code chunk. If you do not, your document will not compile!

#a
```{r eval = TRUE, echo = TRUE, error = TRUE}
# function: abs_val, first checks if input is numeric and returns its absolute 
#           value if it is and print error message if it is not.
# input: x
# output: numeric result equals absolute value of x
abs_val <- function(x) {
    if (is.numeric(x)) {
        if (x >= 0) {
            return(x)
        } else {
            return(-x)
        }
    } else {
        cat("x is not numeric!")
    }
}
```

#b
```{r eval = TRUE, echo = TRUE, error = TRUE}
# evaluating the function
# let input be positive number and check if its absolute value equals itself.
# if equals, the function works
x <- 7
x == abs_val(x)
```

#c
```{r eval = TRUE, echo = TRUE, error = TRUE}
# evaluating the function
# let input be negative number and check if its absolute value equals itself.
# if does not equal, the function works
y <- -7
y == abs_val(y)
```

#d
```{r eval = TRUE, echo = TRUE, error = TRUE}
# evaluating the function
# let input be non-numeric and check if the function works.
# if prints error message, the function works
z <- "string"
abs_val(z)
```

### Part B. Loops

3. Write and demonstrate a function that loops through a numeric input vector
and returns a numeric vector of the same length, where the $i$th element of the output is equal to the sum of the first $i$ elements of the input. (Hint: The first element of the output should match the first element of the input. The last element of the output should match the sum of the elements of the input.)

```{r eval = TRUE, echo = TRUE, error = TRUE}
# function: sum_loop, sum up the first i elements of input and put the sum as 
#           ith element of output
# input: vector x, must be numeric
# output: numeric vector s
sum_loop <- function(x) {
    # check if input is numeric
    if (!is.numeric(x)) {
        cat("input must be numeric")
    } else {
        s <- rep(NA, length(x))
        my_sum <- 0
        for (i in 1:length(x)) {
            my_sum <- my_sum + x[i]
            s[i] <- my_sum
        }
        return(s)
    }
}
```

```{r eval = TRUE, echo = TRUE, error = TRUE}
# evaluating the function
# if function works, return vector (1, 3, 6, 10)
sum_loop(c(1, 2, 3, 4))
```

```{r eval = TRUE, echo = TRUE, error = TRUE}
# evaluating the function
# if function works, print error message "input must be numeric"
sum_loop(c("a", "b"))
```


4. Write and demonstrate a function that loops through each element of a matrix and replaces each element with the row index minus the column index. (Hint: use `ncol()` and `nrow()`)

```{r eval = TRUE, echo = TRUE, error = TRUE}
# function: modify_matrix, check if matrix valid; if so, modify each matrix 
#           element to the row index minus the column index; if not, print
#           error message
# input: matrix x
# output: numeric matrix x
modify_matrix <- function(x) {
    # check if matrix valid
    if (nrow(x) > 0 && ncol(x) > 0){
        for (i in 1:nrow(x)) {
            for (j in 1:ncol(x)) {
                x[i, j] <- i - j
            }
        }
        return(x)
    } else {
        cat("input not valid")
    }
}
```

```{r eval = TRUE, echo = TRUE, error = TRUE}
# evaluating the function
# if function works, return matrix (0 -1 -2; 1 0 -1; 2 1 0; 3 2 1)
m <- matrix(NA, nrow = 4, ncol = 3)
modify_matrix(m)
```

```{r eval = TRUE, echo = TRUE, error = TRUE}
# evaluating the function
# if function works, print error message "input not valid"
n <- matrix(NA, nrow = 0, ncol = 0)
modify_matrix(n)
```









