---
title: "Lab 1"
author: "Sylvia Du"
date: "1/13/20"
output: html_document
---

## Part 1. Vectors


#1a.
```{r, eval = TRUE, echo = TRUE}
# use rnorm to generate a random Norma distribution of 100000 elements and store
# them in a vector. print out the first 5 elements of the vector.
randNorm <- c(rnorm(100000, 7, 3))
head(randNorm, 5)
```
#1b.
The larger the sample size gets, the more likely that the distribution of the 
data on the histogram would be bell-shaped. This is because the more element we
take, the more likely the data would be in Normal distribution as expected and
the less likely that the outliers would effect the resulting distribution.
```{r, eval = TRUE, echo = TRUE}
# generate 5 histograms of the first 10, 100, 1000, 10000, 100000 elements of 
# the vector respectively.
tenth <- head(randNorm, 10)
hist(tenth, main = paste("First 10 elements of vector"), xlab = "value")
hundred <- head(randNorm, 100)
hist(hundred, main = paste("First 100 elements of vector"), xlab = "value")
thousand <- head(randNorm, 1000)
hist(thousand, main = paste("First 1000 elements of vector"), xlab = "value")
tenthou <- head(randNorm, 10000)
hist(tenthou, main = paste("First 10000 elements of vector"), xlab = "value")
hist(randNorm, main = paste("All elements of vector"), xlab = "value")
```

#1c.

```{r, eval = TRUE, echo = TRUE}
# standardize the vector of random elements and store the elements in a new 
# vector. generate a histogram of the standarized vector.
total_sd <- (randNorm - mean(randNorm)) / sd(randNorm)
m_rN <- mean(total_sd)
sd_rN <- sd(total_sd)
hist(total_sd, main = paste("All elements of standarized vector"),
     xlab = "value")
```

The distribution of the histogram is bell-shaped and similar with the histogram
of the original vector but with mean of `r m_rN` (about 0) and standard
deviation of `r sd_rN` (about 1).


#1d.

```{r, eval = TRUE, echo = TRUE}
# calculate the percent of simulations from expected Normal(0, 1) and from my
# randomly generated Normal distribution that are above 1.644854 respectively.
percentile <- pnorm(1.644854, mean = 0, sd = 1, lower.tail = FALSE,
                    log.p = FALSE)
my_percentile <- mean(total_sd > 1.644854)
```

The expected percentile to be above 1.644854 of Normal(0, 1) is `r percentile`.
The observed percentile to be above 1.644854 of my standarized simulations is
`r my_percentile`. My observed percentile is slightly greater than the expected
percentile.

#1e.
```{r, eval = TRUE, echo = TRUE}
# calculate the percent of elements that are above 1.644854 of the first 10, 
# 100, 1000, 10000 elements of the standarized vector respectively.
ten_sd <- mean(head(total_sd, 10) > 1.644854)
hund_sd <- mean(head(total_sd, 100) > 1.644854)
thou_sd <- mean(head(total_sd, 1000) > 1.644854)
tentho_sd <- mean(head(total_sd, 10000) > 1.644854)
```

The percentiles of elements that are above 1.644854 of the first 10, 100, 1000,
10000 elements of the standarized vector are `r ten_sd`, `r hund_sd`, 
`r thou_sd`, `r tentho_sd`. The larger the sample size gets, the closer that the
percentile gets to the expected percentile.

#1f.
```{r, eval = TRUE, echo = TRUE}
# calculate the percentile of the given value in my random simulation of Normal
# distribution. then standarize the given value and calculate the percentile of
# the standarized value in the standarized simulation.
f_percentile <- mean(randNorm < 13.86)
f_standard <- (13.86 - mean(randNorm)) / sd(randNorm)
f_sd_percentile <- mean(total_sd < f_standard)
```
The percentile of 13.86 in my random simulation of Normal distribution
is `r f_percentile` and the percentile of the standarized value `r f_standard`
in my standarized simulation is `r f_sd_percentile`. 
The two values are the same.

#1g.
```{r, eval = TRUE, echo = TRUE}
extreme <- 2 * pnorm(f_standard, lower.tail = FALSE)
```
I expect the percentile of simulations from an expected standard Normal
distribution that is more extreme than the standarized value to be about 2 times
1 minus the the percentile of the standarized value in my standarized 
simulation. The expected percentile is `r extreme`.

#1h.
It is not very likely that the value given was drawn from the same distribution
as my simulations, that is, from Normal(7, 3). The value is more than 2 times 
the standard deviation larger than the mean, so it is very unlikely that this
number would be simulated from N(7, 3).


## Part 2.

#2a.
```{r, eval = TRUE, echo = TRUE}
# create two empty matrices, the smaller one with 100 rows and the larger one
# with 100000 rows, both of the matrices with 4 columns.
my_matrix_s <- matrix(, nrow = 100, ncol = 4)
my_matrix_l <- matrix(, nrow = 100000, ncol = 4)
```

#2b.
```{r, eval = TRUE, echo = TRUE}
# create random Binomial distributions with n = 5 (size) and probability of 0.2,
# 0.4, 0.6, 0.8 respectively, save the simulations in different vectors. fill 
# the matrices with random simulations of different number of trials (size = 100
# for the small matrix and 100000 for the large matrix). print first five rows
# of both matrix.
my_vector_s1 <- c(rbinom(100, 5, 0.2))
my_vector_s2 <- c(rbinom(100, 5, 0.4))
my_vector_s3 <- c(rbinom(100, 5, 0.6))
my_vector_s4 <- c(rbinom(100, 5, 0.8))
my_matrix_s[, 1] <- my_vector_s1
my_matrix_s[, 2] <- my_vector_s2
my_matrix_s[, 3] <- my_vector_s3
my_matrix_s[, 4] <- my_vector_s4

my_vector_l1 <- c(rbinom(100000, 5, 0.2))
my_vector_l2 <- c(rbinom(100000, 5, 0.4))
my_vector_l3 <- c(rbinom(100000, 5, 0.6))
my_vector_l4 <- c(rbinom(100000, 5, 0.8))
my_matrix_l[, 1] <- my_vector_l1
my_matrix_l[, 2] <- my_vector_l2
my_matrix_l[, 3] <- my_vector_l3
my_matrix_l[, 4] <- my_vector_l4

my_matrix_s[1:5, ]
my_matrix_l[1:5, ]
```

#2c.
For both of the two matrices, the larger the probability get, the larger that 
the largest proportion of elements get. In other words, the curves shift to the
right as the probability gets larger. Comparing the two matrices, the larger
matrix which has more elements displays the characteristics better.
```{r, eval = TRUE, echo = TRUE}
# create 8 histograms of the columns respectively.
hist(my_vector_s1, main = paste("First column of small matrix"), 
     xlab = "value", breaks = 11)
hist(my_vector_s2, main = paste("Second column of small matrix"), 
     xlab = "value", breaks = 11)
hist(my_vector_s3, main = paste("Third column of small matrix"), 
     xlab = "value", breaks = 11)
hist(my_vector_s4, main = paste("Fourth column of small matrix"),
     xlab = "value", breaks = 11)

hist(my_vector_l1, main = paste("First column of large matrix"), 
     xlab = "value", breaks = 11)
hist(my_vector_l2, main = paste("Second column of large matrix"), 
     xlab = "value", breaks = 11)
hist(my_vector_l3, main = paste("Third column of large matrix"), 
     xlab = "value", breaks = 11)
hist(my_vector_l4, main = paste("Fourth column of large matrix"),
     xlab = "value", breaks = 11)
```

#2d.
```{r, eval = TRUE, echo = TRUE}
# calculate the mean value of each column for both matrices, and put the results
# in a table.
mean_s1 <- mean(my_vector_s1)
mean_s2 <- mean(my_vector_s2)
mean_s3 <- mean(my_vector_s3)
mean_s4 <- mean(my_vector_s4)
mean_l1 <- mean(my_vector_l1)
mean_l2 <- mean(my_vector_l2)
mean_l3 <- mean(my_vector_l3)
mean_l4 <- mean(my_vector_l4)
mean_table <- matrix(c(mean_s1, mean_s2, mean_s3, mean_s4, mean_l1, mean_l2,
                       mean_l3, mean_l4), ncol=2, byrow = FALSE)
rownames(mean_table) <- c("First column", "Second column", "Third column", 
                          "Fourth column")
colnames(mean_table) <- c("Matrix of 100 rows", "Matrix of 100000 rows")
mean_table <- as.table(mean_table)
mean_table
```

#2e.
The expected column mean of each column should be n (size) times probability, 
which would be 1, 2, 3, 4 respectively for each columns of both matrices. The
larger matrix has observed column means that are closer to the expectation. 
This is because that the larger the sample size gets, the less that the outliers
would effect the whole data set.

#2f.
Compared to the trial of n = 5, the characteristics that the curves shift to the
right as the probability gets larger remains unchanged. Comparing the modified
two matrices, the larger modified matrix display this characteristics clearer 
than the smaller modified matrix.
```{r, eval = TRUE, echo = TRUE}
# create two empty matrices, the smaller one with 100 rows and the larger one
# with 100000 rows, both of the matrices with 4 columns.
# create random Binomial distributions with n = 1000 (size) and probability of 
# 0.2, 0.4, 0.6, 0.8 respectively, save the simulations in different vectors. 
# fill the matrices with random simulations of different number of trials 
# (size = 100 for the small matrix and 100000 for the large matrix). 
# print first five rows.
# of both matrix.
# create 8 histograms of the columns respectively.
# calculate the mean value of each column for both matrices, and put the results
# in a table.

mod_matrix_s <- matrix(1:400, nrow = 100, ncol = 4)
mod_matrix_l <- matrix(1:400000, nrow = 100000, ncol = 4)

mod_vector_s1 <- c(rbinom(100, 1000, 0.2))
mod_vector_s2 <- c(rbinom(100, 1000, 0.4))
mod_vector_s3 <- c(rbinom(100, 1000, 0.6))
mod_vector_s4 <- c(rbinom(100, 1000, 0.8))
mod_matrix_s[, 1] <- mod_vector_s1
mod_matrix_s[, 2] <- mod_vector_s2
mod_matrix_s[, 3] <- mod_vector_s3
mod_matrix_s[, 4] <- mod_vector_s4

mod_vector_l1 <- c(rbinom(100000, 1000, 0.2))
mod_vector_l2 <- c(rbinom(100000, 1000, 0.4))
mod_vector_l3 <- c(rbinom(100000, 1000, 0.6))
mod_vector_l4 <- c(rbinom(100000, 1000, 0.8))
mod_matrix_l[, 1] <- mod_vector_l1
mod_matrix_l[, 2] <- mod_vector_l2
mod_matrix_l[, 3] <- mod_vector_l3
mod_matrix_l[, 4] <- mod_vector_l4

mod_matrix_s[1:5, ]
mod_matrix_l[1:5, ]

hist(mod_vector_s1, main = paste("First column of small matrix"), 
     xlab = "value", breaks = 110)
hist(mod_vector_s2, main = paste("Second column of small matrix"), 
     xlab = "value", breaks = 110)
hist(mod_vector_s3, main = paste("Third column of small matrix"), 
     xlab = "value", breaks = 110)
hist(mod_vector_s4, main = paste("Fourth column of small matrix"),
     xlab = "value", breaks = 110)

hist(mod_vector_l1, main = paste("First column of large matrix"), 
     xlab = "value", breaks = 110)
hist(mod_vector_l2, main = paste("Second column of large matrix"), 
     xlab = "value", breaks = 110)
hist(mod_vector_l3, main = paste("Third column of large matrix"), 
     xlab = "value", breaks = 110)
hist(mod_vector_l4, main = paste("Fourth column of large matrix"),
     xlab = "value", breaks = 110)

mod_mean_s1 <- mean(mod_vector_s1)
mod_mean_s2 <- mean(mod_vector_s2)
mod_mean_s3 <- mean(mod_vector_s3)
mod_mean_s4 <- mean(mod_vector_s4)
mod_mean_l1 <- mean(mod_vector_l1)
mod_mean_l2 <- mean(mod_vector_l2)
mod_mean_l3 <- mean(mod_vector_l3)
mod_mean_l4 <- mean(mod_vector_l4)
mod_mean_table <- matrix(c(mod_mean_s1, mod_mean_s2, mod_mean_s3, mod_mean_s4, 
                           mod_mean_l1, mod_mean_l2, mod_mean_l3, mod_mean_l4),
                         ncol=2, byrow = FALSE)
rownames(mod_mean_table) <- c("First column", "Second column", "Third column", 
                          "Fourth column")
colnames(mod_mean_table) <- c("Matrix of 100 rows", "Matrix of 100000 rows")
mod_mean_table <- as.table(mod_mean_table)
```

#2g.
The larger matrix has more extreme maxs and mins. The reason should be that the 
larger the sample size is, the chances of generating extreme values would be 
greater.
```{r, eval = TRUE, echo = TRUE}
# create five-number summaries of each column in both matrices (with summary 
# method but excluding the mean value). create a table of the 5-number summary.
sum_s1 <- summary(mod_vector_s1)[-4]
sum_s2 <- summary(mod_vector_s2)[-4]
sum_s3 <- summary(mod_vector_s3)[-4]
sum_s4 <- summary(mod_vector_s4)[-4]
sum_l1 <- summary(mod_vector_l1)[-4]
sum_l2 <- summary(mod_vector_l2)[-4]
sum_l3 <- summary(mod_vector_l3)[-4]
sum_l4 <- summary(mod_vector_l4)[-4]

sum_table <- rbind(sum_s1, sum_s2, sum_s3, sum_s4, sum_l1, sum_l2, sum_l3,
                   sum_l4)
colnames(sum_table) <- c("Min", "First Quantile", "Median", "Third Quantile", 
                         "Max")
rownames(sum_table) <- c("First col of small matrix", 
                         "Second col of small matrix", 
                         "Third col of small matrix", 
                         "Fourth col of small matrix", 
                         "First col of large matrix", 
                         "Second col of large matrix", 
                         "Third col of large matrix", 
                         "Fourth col of large matrix")
sum_table <- as.table(sum_table)
sum_table
```




